A More Complete Algebra for File Synchronization

### 2015-11-18 2015/paper.tex
Proof completed

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}

\newcommand{\works}[1]{\dot{\mathbf{w}}({#1})}
\newcommand{\worksc}[2]{\dot{\mathbf{w}}({#1}|{#2})}

@@

\newcommand{\infer}[2]{\begin{array}{c}{#1}\\\hline{#2}\end{array}}
\newcommand{\inferr}[3]{\begin{array}{c}{#1}\\{#2}\\\hline{#3}\end{array}}

@@

\begin{mydef}
Using $\FS\{p\mapsto V\}$ we denote a filesystem derived from $\FS$:
\[ \FS\{p\mapsto V\}(q) :=
   \begin{cases}
   V &\mbox{if~} q=p\\
   \FS(q) &\mbox{otherwise.}
   \end{cases}
\]
\end{mydef}

The exact behaviour of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxy(p,V_Y)$ where
$X$ and $Y$ are types ($\fsregcommandchar{D}$, $\fsregcommandchar{F}$ or $\fsregcommandchar{\empt}$),
$p\in\setp$ and $V_Y$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $V_Y$ if there is only one suitable value.
\[ \cxy(p,V_Y)\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(p)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS\{p\mapsto V_Y\} \mbox{~violates the tree property}\\
   \FS\{p\mapsto V_Y\} &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

\begin{mydef}{$p\descendant q$.}
$p\descendant q$ if $p$ is the ancestor of $q$ ($p=\parent^n(q)$).
\end{mydef}

@@

\begin{mydef}
$A\equiv B$ means that $A$ and $B$ are equivalent.
Its intended interpretation is that they behave in the same way on
all filesystems: $\forall \FS: A\FS=B\FS$.
It is a commutative and transitive relation.
\end{mydef}

\begin{mydef}
$A\eqext B$ means that $B$ extends $A$.
Its intended interpretation is that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS: A\FS\neq\fsbroken\Rightarrow A\FS=B\FS$
It is a transitive relation.
\end{mydef}

\begin{mydef}
$\works{A_1,A_2,\ldots,A_n}$ means that 
$A_1,A_2,\ldots,A_n$ work at the same time.
Its intended interpretation is
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_n\FS\neq\fsbroken\]
\end{mydef}

\begin{mydef}
$\worksc{A_1,A_2,\ldots,A_n}{B_1,B_2,\ldots,B_n}$ means that 
$A_1,A_2,\ldots,A_n$ work where $B_1,B_2,\ldots,B_n$ work.
Its intended interpretation is
\begin{align*}
\forall \FS:&\\ 
&B_1\FS\neq\fsbroken \wedge \cdots \wedge B_n\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_n\FS\neq\fsbroken
\end{align*}
\end{mydef}

@@

\begin{myrul}\label{r_ex_w}
\[\infer{A\eqnrw B}{\worksc{A}{B}}\]
\end{myrul}

@@

\begin{myrul}
\[\infer{\worksc{A_1,\ldots,A_i,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{A_i,A_1,\ldots,A_{i-1},A_{i+1},\ldots,A_n}{B_1,\ldots,B_n}}\]
\end{myrul}

@@

@@

@@

\begin{myax}\label{ax_separate_commute}
Commands on unrelated paths commute:
$\cxy(p,V_Y); \czw(q,V_W) \equiv \czw(q,V_W); \cxy(p,V_Y)$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated paths do not break every filesystem:
$\cxy(p,V_Y); \czw(q,V_W) \nequiv \cbrk$ where $p\not\descendant q$ and $q\not\descendant p$.
\end{myax}

@@

\begin{myax}\label{ax_distantparent_breaks}
$\cxy(p,V_Y); \czw(\pp,V_W) \equiv \cbrk$ where $p\ancestor \pp$ and $\parent(p)\neq \pp$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$.
\end{myax}

@@

\begin{mydef}{Reconciliation}\label{def:reconciliation}
\begin{align*}
\recaset = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: [c;c']\equiv[c';c]\\
\} & \\
\reca = \ordered{\recaset}&
\end{align*}

In other words, the sequence $\reca$ we apply to $\FS_B$ consists of all commands from $A\setminus B$
that commute with every command in $B\setminus A$.
\end{mydef}

@@

\begin{myth}
If $A, B$ are simple, then $\worksc{[B;\reca]}{A,B}$
\end{myth}

@@

@@

\begin{mylem}
If $S\coworks T$ and $\forall s\in S, t\in T: [s;t]\equiv[t;s]$
then $[S;T]\coworks S$.
In this case we also know
$\forall \FS: S\FS\neq\fsbroken \wedge T\FS\neq\fsbroken \Rightarrow [S;T]\FS\neq\fsbroken$
(but we can't yet express this in the algebra).
\end{mylem}

\begin{mylem}
$[S;T]\coworks[S;U]\Rightarrow T\coworks U$.
TODO Is this an axiom?
\end{mylem}

Turning back to reconciliation we know
$A\equiv[\ordered{A\cap B};A_0]$
where $A_0=\ordered{A\setminus B}$.
We define $B_0$ similarly.
As $A\coworks B$ we know $A_0\coworks B_0$.

We also know commands in $A_0$ and $\recb$ commute, that is,
$\forall a\in A_0, b\in\recb: [a;b]\equiv[b;a]$.

>>>>

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyx}{\fscommand{Y}{X}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\cxw}{\fscommand{X}{W}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

@@

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\indep}{\not\sim}

\newcommand{\works}[1]{{\mathbf{w}}({#1})}
\newcommand{\worksc}[2]{{\mathbf{w}}({#1}|{#2})}

@@

\newcommand{\infer}[2]{\begin{array}{c}{#1}\\\hline{#2}\end{array}}
\newcommand{\inferr}[3]{\begin{array}{c}{#1}\\{#2}\\\hline{#3}\end{array}}
\newcommand{\inferrr}[4]{\begin{array}{c}{#1}\\{#2}\\{#3}\\\hline{#4}\end{array}}

@@

\begin{mydef}
Using $\FS[V/p]$ we denote a filesystem derived from $\FS$:
\[ \FS\{p\mapsto V\}(q) :=
   \begin{cases}
   V &\mbox{if~} q=p\\
   \FS(q) &\mbox{otherwise.}
   \end{cases}
\]
\end{mydef}

The exact behaviour of our commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxy(p,V_Y)$ where
$X$ and $Y$ are types ($\fsregcommandchar{D}$, $\fsregcommandchar{F}$ or $\fsregcommandchar{\empt}$),
$p\in\setp$ and $V_Y$ is a value of the appropriate type from $\setvx{Y}$.
At times we omit $V_Y$ if there is only one suitable value.
\[ \cxy(p,V_Y)\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(p)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/p] \mbox{~violates the tree property}\\
   \FS[V/p] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

\begin{mydef}{$p\descendant q$.}
$p\descendant q$ if $p$ is the ancestor of $q$ ($p=\parent^n(q)$). (the transitive closure of the $\parent$ relation)
\end{mydef}

@@

\begin{mydef}
$A\equiv B$ means that $A$ and $B$ are equivalent.
We assume that it is a reflexive, commutative and transitive relation.
Its intended interpretation is that they behave in the same way on
all filesystems: $\forall \FS: A\FS=B\FS$.
\end{mydef}

\begin{mydef}
$A\eqext B$ means that $B$ extends $A$.
It is a reflexive and transitive relation.
Its intended interpretation is that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS: A\FS\neq\fsbroken\Rightarrow A\FS=B\FS$
\end{mydef}

\begin{mydef}
$\works{A_1,A_2,\ldots,A_n}$ means that 
$A_1,A_2,\ldots,A_n$ work at the same time.
TODO: $A_i$ is a set (order not important).
Its intended interpretation is
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_n\FS\neq\fsbroken\]
\end{mydef}

\begin{mydef}
$\worksc{A_1,A_2,\ldots,A_n}{B_1,B_2,\ldots,B_n}$ means that 
all of $A_1,A_2,\ldots,A_n$ work where all of $B_1,B_2,\ldots,B_n$ work.
TODO: $A$s $B$s are sets.
We assume that the arguments $A_i$ and $B_i$ commute among each other,
and that the relation is reflexive, that is, $\worksc{B}{B}$ is true for all sequences.
Its intended interpretation is
\begin{align*}
\forall \FS:&\\ 
&B_1\FS\neq\fsbroken \wedge \cdots \wedge B_n\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_n\FS\neq\fsbroken
\end{align*}
\end{mydef}

@@

\begin{myrul}\label{r_ex_w}
\[\infer{B\eqext A}{\worksc{A}{B}}\]
\end{myrul}

@@

% \begin{myrul}
% \[\infer{\worksc{A_1,\ldots,A_i,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{A_i,A_1,\ldots,A_{i-1},A_{i+1},\ldots,A_n}{B_1,\ldots,B_n}}\]
% \end{myrul}

\begin{myrul}\label{r_w_combine}
\[\inferr{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{C}{B_1,\ldots,B_n}}{\worksc{A_1,\ldots,A_n,C}{B_1,\ldots,B_n}}\]
\end{myrul}

@@

% \begin{myrul}\label{r_w_replacecond}
% \[\inferr{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_n}}{\worksc{C}{B_n}}{\worksc{A_1,\ldots,A_n}{B_1,\ldots,B_{n-1},C}}\]
% \end{myrul}

@@

\begin{mydef}{Inverse commands and sequences.}
The inverse of command $\cxy(p,V_Y)$ is $\cxy(p,V_Y)^{-1} = \cyx(p,V_X)$
where $V_X$ is an arbitrary value from $\setvx{X}$.
The inverse of sequence $S$, $S^{-1}$, consists of the inverses of the commands in $S$
in reverse order.
\end{mydef}

\begin{myrul}\label{r_invmove}
\[ \infer{\worksc{S;A}{B}}{\worksc{A}{S^{-1};B}} \]
To see the soundness of this rule...
Sequences are bijections on filesystems over type-equality.
TODO: $A,S,B$ need to be simple?
\end{myrul}

\begin{mydef}{Independent commands, sequences and sets.}
For two commands $\cxy$ and $\czw$ we write $\cxy\indep\czw$ if $\cxy;\czw\equiv\czw;\cxy\nequiv\cbrk$.
% TODO use w() instead of nequiv break
For two sequences or sets of commands $A$ and $B$ we write $A\indep B$ if
$\forall a\in A \mbox{and~} \forall b\in B: a\indep b$.
\end{mydef}

\begin{myrul}\label{r_indep_combine}
\[ \inferrr{\worksc{T}{S}}{\worksc{U}{S}}{T\indep U}{\worksc{T;U}{S}} \]
To see the soundness of this rule...
\end{myrul}

\begin{mylem}\label{indep_prefix_combine}
\[ \inferrr{\worksc{A;B}{S}}{\worksc{A;C}{S}}{B\indep C}{\worksc{A;B;C}{S}} \]
\end{mylem}
\begin{proof}
From \cref{r_invmove,r_indep_combine}
TODO reverse of /cref{r_invmove} needed
\end{proof}

@@

\begin{myax}\label{ax_separate_commute}
Commands on unrelated paths commute:
$\cxy(p,V_Y); \czw(q,V_W) \equiv \czw(q,V_W); \cxy(p,V_Y)$ where $p\not\descendant q$ and $q\not\descendant p$.
TODO $p\neq q$
\end{myax}

\begin{myax}\label{ax_separate_nobreaks}
Commands on unrelated paths do not break every filesystem:
$\cxy(p,V_Y); \czw(q,V_W) \nequiv \cbrk$ where $p\not\descendant q$ and $q\not\descendant p$.
TODO $p\neq q$
\end{myax}

% TODO Use 'node' instead of 'path'

@@

\begin{myax}\label{ax_distantparent_breaks}
$\cxy(p,V_Y); \czw(\pp,V_W) \equiv \cbrk$ where $p\ancestor \pp$ and $\parent(p)\neq \pp$ and $\cxy\neq\cbb$ and $\czw\neq\cdd$.
TODO merge into previous
\end{myax}

@@

\begin{mydef}{Reconciliation}\label{def:reconciliation}
\begin{align*}
\recaset = \{c| & \\
& c\in A\setminus B \\
& \wedge \\
& \forall c'\in B\setminus A: c\indep c'\\
\} & \\
\reca = \ordered{\recaset}&
\end{align*}

In other words, the sequence $\reca$ we apply to $\FS_B$ consists of all commands from $A\setminus B$
that independent from every command in $B\setminus A$.
\end{mydef}

@@

\begin{myth}
If $A, B$ are simple, then $\worksc{B;\reca}{A,B}$
\end{myth}

@@

TODO: unmarked-marked can be reversed -- but induction is wrong

@@

We would like to prove that $\worksc{B;\reca}{A,B}$.
From above we know that this is equivalent to
$\worksc{\ordered{A\cap B};\ordered{B\setminus A};\reca}{A,B}$.
As we already know that $\reca\indep B\setminus A$
and that $\worksc{\ordered{A\cap B};\ordered{B\setminus A}}{A,B}$
(as $\worksc{B}{A,B}$),
based on \cref{indep_prefix_combine}
we only need to prove that
$\worksc{\ordered{A\cap B};\reca}{A,B}$.

\newcommand{\acbi}{\ordered{A\cap B}^{-1}}
\newcommand{\amb}{A\setminus B}
\newcommand{\bma}{B\setminus A}
Using \cref{r_invmove} we can rephrase this as
% TODO axiom doesn't specify multiple conditions
$\worksc{\reca}{\acbi;A,\acbi;B}$.

We already know the following things:
\begin{itemize}
\item $\amb$ and $\bma$ are simple as $A$ and $B$ are simple
% \item $\amb \cap \bma = \emptyset$
\item $\reca$ is the largest subset of $\amb$ for which $\reca\indep\amb$
\item $\worksc{\ordered{\amb};\ordered{\bma}}{\acbi;A,\acbi;B}$ as $\worksc{A,B}{A,B}$ based on \cref{r_invmove}
\end{itemize}

Our theorem is therefore proven if we prove the following lemma:

\begin{mylem}
If
   \begin{itemize}
   \item \newcommand{\condSimple}{(1)} \condSimple $S$ and $T$ are simple sequences
   % \item \newcommand{\condDisj}{(4)} \condDisj $S\cap T=\emptyset$
   \item \newcommand{\condApr}{(2)} \condApr $S^*$ is the largest subset of $S$ where $S^*\indep T$
   \item \newcommand{\condWork}{(3)} \condWork $\worksc{S,T}{C_1,\ldots,C_n}$
   \end{itemize}
then
\[ \worksc{S^*}{C_1,\ldots,C_n} \]
\end{mylem}

\begin{proof}
% TODO Refer back to conditions!
This is trivial unless $\works{C_1,\ldots,C_n}$, so we will assume that it is true
and therefore $\works{S,T}$.
% TODO This probably needs an axiom

Similarly to above, we prove that $S$ can be transformed into $S'$ by equivalences
so that all commands in $S^*$ would be at its beginning.
If so then based on \cref{worksextpostfix}
$\worksc{S^*}{C_1,\ldots,C_n}$ as $\worksc{S'}{C_1,\ldots,C_n}$.

Also similarly to above,
we proceed by induction on the number of commands in $S$ that are not in $S^*$, but which precede
the last command in $S$ that is also in $S^*$.
If the number is 0, the lemma is true.
Otherwise let the last command in $S$ also in $S^*$ be $\cxy(p,V_Y)$
and the preceding command in $S$ be $\czw(q,V_W)$.

As $S$ is simple and $\works{S}$, from 
\cref{ax_distantchild_breaks,ax_distantparent_breaks,ax_directchild_breaks,ax_directparent_breaks}
we know that these commands can only be on unrelated paths or form a construction or destruction pair.
In the first case, swapping the commands results in a sequence equivalent to $S$,
with the number on relevant commands not in $S^*$ reduced by one.
In the last two cases,
either $q=\parent(p)$ or $p=\parent(q)$.

We also know that there must be a command $\cqr(r,V_R)$ in $T$ which is not independent of $\czw(q,V_W)$.
By definition this means that one of the following must be true:
\begin{itemize}
\item $\czw(q,V_W);\cqr(r,V_R) \equiv\cbrk$
\item $\cqr(r,V_R);\czw(q,V_W) \equiv\cbrk$
\item $\czw(q,V_W);\cqr(r,V_R) \nequiv \cqr(r,V_R);\czw(q,V_W)$
\end{itemize}
From these and \cref{ax_separate_commute,ax_separate_nobreaks} we also know that
either $q\descendant r$ or $r\descendant q$.
By definition we also know that $\cxy(p,V_Y)\indep \cqr(r,V_R)$,
and the same argument shows that therefore
$p\not\descendant r$ and $r\not\descendant p$ and $p\neq r$. TODO CHECK

We therefore have four cases considering the relationships between $p,q$ and $r$:
\begin{itemize}
\item $p=\parent(q) \wedge r\descendant q$.
   This would mean that $r\descendant p$ or $r=p$, which is a contradiction.
\item $p=\parent(q) \wedge q\descendant r$.
   This would mean that $p\descendant r$, which is a contradiction.
\item $q=\parent(p) \wedge r\descendant q$.
   This would mean that $r\descendant p$, which is a contradiction.  
\item $q=\parent(p) \wedge q\descendant r$.
   Let us continue with this case.
\end{itemize}

We know that $T$ cannot have a command on $q$ as $q=\parent(p)$ and $\cxy(p,V_Y)\indep T$.
Create $T'\equiv T$ by inserting a command on $q$ before $\cqr(r,V_R)$
into $T$ according to \cref{ax_parent_assert}.
We now have $\cxy(p,V_Y)\indep T\equiv T'$ which is a contradiction as
$T'$ has a command on $q$ and $q=\parent(p)$.
% TODO axiom? A \~ B == C  =>  A \~ C
\end{proof}

TODO $\reca$ is maximal
